\begin{thebibliography}{9}

\bibitem{amdahl}{\sc G.M. Amdahl},
{\em Validity of the Single-processor Approach to Achieving
Large-scale Computing Capabilities}, In {\em AFIPS Conference
Proceddings} {\bf 30} (Atlantic City, N.J., April 18-20), AFIPS 
Press (1967).

\bibitem{christo}{\sc E. Balas and P. Toth},
{\em Branch and Bound Methods}, in E.L. Lawler, J.K. Lenstra,
A.H.G. Rinnooy Kan and D.B. Shmoys, eds., The Traveling Salesman
Problem: A Guided Tour of Combinatorial Optimization, Wiley, New York
(1985), 361.

\bibitem{BB&G}{\sc R.L. Boehning R.M. Butler and B.E. Gillet}, {\em A
Parallel Integer Linear Programming Algorithm}, European Journal of
Operations Research {\bf 34} (1988).

\bibitem{chvatal}{\sc V. Chv\'atal}, 
{\em Linear Programming},
W.H. Freeman and Company (1983).

\bibitem{cplex} 
{\em Using the CPLEX$^{\copyright}$ Callable Library},
CPLEX$^{\copyright}$ Optimization Inc. (1994).

\bibitem{C&P}{\sc H. Crowder and M. Padberg}, 
{\em Solving Large Scale Symmetric Traveling Salesman Problems to
Optimality}, Management Science {\bf 26} (1980), 495.

\bibitem{large01}{\sc H. Crowder, E.L. Johnson and M. Padberg}, 
{\em Solving Large-Scale Zero-One Linear Programming Problems},
Operations Research {\bf 31} (1983), 803.

\bibitem{CDCMR}{\sc V.-D. Cung, S. Dowaji, B. Le Cun, T. Mauthor and
C. Roucairol}, {\em Concurrent Data Structures and Load Balancing
Strategies for Parallel Branch and Bound/A* Algorithms}, DIMACS Series
in Discrete Optimization and Theoretical Computer Science {\bf 30} (1997).

\bibitem{eckstein}{\sc J. Eckstein}, {\em Parallel Branch and Bound
Algorithms for General Mixed Integer Programming on the CM-5}, SIAM
Journal on Optimization {\bf 4} (1994).

\bibitem{eckstein2}{\sc J. Eckstein}, {\em How Much Communication Does
Parallel Branch and Bound Need?}, INFORMS
Journal on Computing {\bf 9} (1997).

\bibitem{Marta}{\sc M. Eso},
{\em Parallel Branch and Cut for Set Partitioning},
Ph.D. Dissertation, Field of Operations Research,
Cornell University, Ithaca, N.Y., U.S.A. (1999).

\bibitem{G&J}{\sc M.R. Garey and D.S. Johnson}, 
{\em Computers and Intractability: A Guide to the Theory of
NP-Completeness},
W.H. Freeman and Co., San Francisco (1979).                

\bibitem{G&N}{\sc R.S. Garfinkel and G.L. Nemhauser},
{\em Integer Programming},
Wiley, New York (1972).

\bibitem{pvm}{\sc A. Geist et al},
{\em PVM: Parallel Virtual Machine, A User's Guide and Tutorial for
Networked Parallel Computing},
MIT Press, Cambridge, MA (1994).

\bibitem{G&C}{\sc B. Gendron and T.G. Crainic},
{\em Parallel Branch and Bound Algorithms: Survey and Synthesis},
Operations Research {\bf 42} (1994), 1042. 

\bibitem{G&VL}{\sc G.H. Golub and C.F. Van Loan},
{\em Matrix Computations},
The Johns Hopkins University Press, Baltimore (1989).

\bibitem{G&K}{\sc A. Grama and V. Kumar}, {\em Parallel Search
Algorithms for Discrete Optimization Problems}, ORSA Journal on
Computing {\bf 7} (1995).

\bibitem{gustafson}{\sc J.L. Gustafson},
{\em Reevaluating Amdahl's Law}, Communications of the ACM {\bf 31} (1988).

\bibitem{lpcombprob}{\sc K. Hoffman and M. Padberg}, 
{\em LP-Based Combinatorial Problem Solving},
Annals of Operations Research {\bf 4} (1985/86), 145.

\bibitem{kozen}{\sc D.C. Kozen},
{\em The Design and Analysis of Algorithms},
Springer-Verlag, New York (1992).

\bibitem{Leo}{\sc L. Kopman},
{\em A New Generic Separation Algorithm and Its Application to the
Vehicle Routing Problem},
Ph.D. Dissertation, Field of Operations Research,
Cornell University, Ithaca, N.Y., U.S.A. (1999).

\bibitem{K&R}{\sc V. Kumar and V.N. Rao}, {\em Parallel Depth-first
Search. Part II. Analysis.}, International Journal of Parallel
Programming {\bf 16} (1987).

\bibitem{K&G}{\sc V. Kumar and A. Gupta}, {\em Analyzing Scalability
of Parallel Algorithms and Architectures}, Journal of Parallel and
Dsitributed Computing {\bf 22} (1994).

\bibitem{Lthesis}{\sc L. Lad\'{a}nyi},
{\em Parallel Branch and Cut and Its Application to the
       Traveling Salesman Problem},
Ph.D. Dissertation, Field of Operations Research,
Cornell University, Ithaca, N.Y., U.S.A. (1996).

\bibitem{Laursen}{\sc P. Laursen}, {\em Can Parallel Branch and Bound
without Communication Be Effective?}, {\bf 4} (1994).

\bibitem{tspbook}{\sc E.L. Lawler, J.K. Lenstra, A.H.G. Rinnooy
Kan and D.B. Shmoys},
{\em The Traveling Salesman Problem: A Guided
Tour of Combinatorial Optimization},
Wiley, New York (1985).

\bibitem{marsten}{\sc R. Marsten},
{\em The Design of The XMP Linear Programming Library},
ACM Transactions on Mathematical Software {\bf 7} (1981), 481. 

\bibitem{N&W}{\sc G.L. Nemhauser and L.A. Wolsey}, 
{\em Integer and Combinatorial Optimization},
John Wiley \& Sons, Inc. (1988).

\bibitem{P&R}{\sc M. Padberg and G. Rinaldi}, 
{\em A Branch-and-Cut Algorithm for the Resolution of Large-Scale
Traveling Salesman Problems},
SIAM Review {\bf 33} (1991), 60.

\bibitem{vrp-paper}{\sc T.K. Ralphs, L. Kopman, W.R.
Pulleyblank, and L.E. Trotter Jr.},
{\em On the Capacitated Vehicle Rotuing Problem},
submitted for publication.

\bibitem{thesis}{\sc T.K. Ralphs}, 
{\em Parallel Branch and Cut for Vehicle Routing},
Ph.D. Dissertation, Field of Operations Research,
Cornell University, Ithaca, N.Y., U.S.A. (1995).

\bibitem{user-guide}{\sc T.K. Ralphs},
{\em SYMPHONY Version 2.8 User's Guide}, available at 
{\tt www.branchandcut.org/SYMPHONY}.

\bibitem{symphony-future}{\sc T.K. Ralphs and L. Lad\'{a}nyi},
{\em A Generic Parallel Framework for Branch and Cut},
manuscript in preparation.

\bibitem{symphony}{\sc T.K. Ralphs and L. Lad\'{a}nyi},
{\em SYMPHONY: A Parallel Framework for Branch and Cut},
White paper, Rice University (1999).

\bibitem{R&K}{\sc V.N. Rao and V. Kumar}, {\em Parallel Depth-first
Search. Part I. Implementation.}, International Journal of Parallel
Programming {\bf 16} (1987).

\bibitem{R&N}{\sc R. Rushmeier and G. Nemhauser}, {\em Experiments
with Parallel Branch and Bound Algorithms for the Set Covering
Problem}, Operations Research Letters {\bf 13} (1993).

\bibitem{savass}{\sc M.W.P. Savelsbergh}, 
{\em A Branch-and-Price Algorithm for the Generalized
Assignment Problem},
Report COC-9302, Georgia Institute of Technology, Atlanta, Georgia
(1993).

\bibitem{schrijver}{\sc A. Schrijver}, 
{\em Theory of Linear and Integer Programming},
John Wiley \& Sons (1986).

\end{thebibliography}

